---
title       : A simple app to study mortality data
subtitle    : A shiny showcase
author      : Albert de Roos
job         : Coursera Data Science Student
framework   : io2012      # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Background and Rationale

### Mortality data
- the World Health Organization publishes mortality data for a large number of countries over the last 50 years
- this mortality data gives insight in the causes of death and the differences in regions leading to better insight in the causes of disease
- a good graphical presentation of this dat can help in visualizing mortality trends

### A shiny app for mortality data
- we developed a proof-of-concept for visualing the mortality trends using a shiny app
- here we show the R code and shiny server and ui code that was used
- we show a simple app where the user can select the year and the mortality-age group distrbution is show

--- .class #id 

```{r imports, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
graphdata <- readRDS(file="mortdata.Rda") 

```

## Shiny ui.R

- We used a simple layout with header, sidebar and main panels
- We used a simple slider to indicate the year to display
- Code is abbreviated and more text is shown in the app itself

```{r code, eval=FALSE}
shinyUI(pageWithSidebar(
     headerPanel('Pharynx Cancer Mortality Data From Hungary'),
     sidebarPanel(
          p("Use the slider to see the different mortality rate distribution"),
          sliderInput("year", "Choose the year:", min=1955, max=2012, value=1995, sep = "")),
     mainPanel(
          p("Explanatory text here"),
          plotOutput('plot1')
     )
))
```

--- .class #id 

## Shiny server.R and global.R code
- Server code used a ggplot2 grap based on the input year
- We needed an extra file global.R to download the data (not shown)
- The original mortality files contained data on pharynx cancer per year and per age group.
- The input data was melted using dplyr to get tidy data as input for the app.

```{r server, eval=FALSE}

shinyServer(function(input, output, session) {

     output$plot1 <- renderPlot({
          graph <- graphdata[graphdata$Year == input$year, 2:3]
          ggplot(data=graph, aes(x=AgeGroup, y=MortRate, group = 1)) +
               ggtitle("Hungary: Mortality Rates by Age Group") + xlab("Age Group") + 
               ylab("Mortality Rate") +
               geom_point() +
               geom_line() +
               theme(axis.text.x = element_text(angle = 45)) +
               ylim(0,120)
     })
})
```

--- .class #id 

## The Result: The Mortality Data App
- App can be seen live here https://albertderoos.shinyapps.io/MortApp/
- Example graph below for the year 1955 and 1997 

```{r graph_output`, echo=FALSE}
library(ggplot2)
graph <- graphdata[graphdata$Year == 1997, 2:3]
          ggplot(data=graph, aes(x=AgeGroup, y=MortRate, group = 1)) +
               ggtitle("Hungary: Mortality Rates by Age Group") + xlab("Age Group") +
               ylab("Mortality Rate") +
               geom_point() +
               geom_line() +
               theme(axis.text.x = element_text(angle = 45)) +
               ylim(0,120)

```




